---
title: "Final Project: Descriptive Statistics"
author: "Priya Rathod"
date: "Fall 2024"
output: html_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

```{r,  message = FALSE, echo=FALSE, warning=FALSE}

#Loading packages

library(tidyverse) 
library(lubridate)
library(here)
here()
library(dplyr)
library(knitr)

#Importing our dataset on polar bears
PolarBearHair <- read.csv(file = here("Data/Raw/polarBear_CNisotopesHair_beaufortChukchi_rode_1983_2017.csv"), stringsAsFactors = TRUE)

```


```{r,  message = FALSE, echo=FALSE, warning=FALSE}

#Linear Regressions - Correlating Summer Habitats and Nitrogen Isotopes

polarbearhair_hab <- PolarBearHair %>%
  filter(TypicalSummerHabitat %in% c("ice", "land")) 

polarbearhair_hab <- polarbearhair_hab %>%
  mutate(Habitat = ifelse(TypicalSummerHabitat == "ice", 1, 0))

polarbearhair_hab_clean <- 
  select(polarbearhair_hab, TypicalSummerHabitat, Habitat, d15N.Air)

nitrogen_regression <- lm(d15N.Air ~ Habitat, data = polarbearhair_hab_clean)
summary(nitrogen_regression)

```

Through this linear regression model, we wanted to test whether there is any effect of polar bear typical habitat (on ice vs. land) on the polar bear diet
(via the ratio of stable nitrogen isotopes). The output of this regression
suggests that there is a statistically significant relationship between habitat and diet (p-value of 2.04E-07 < 0.01). Given that our dataset had a relatively small sample of habitat data, this analysis should be repeated with a larger dataset.


```{r,  message = FALSE, echo=FALSE, warning=FALSE}

# Plot 1 - Correlating Time and Nitrogen Isotopes

plot1 <- ggplot(PolarBearHair, aes(x = Year, y = d15N.Air, color = Sex)) + geom_smooth(method=lm) + scale_x_continuous(limits = c(2004, 2020), breaks = seq(1985, 2020, by = 5)) + labs(title = "Figure 1 - Stable Nitrogen Isotopes sampled from Polar Bear Hair (2004-2018)",
x = "Year", y = "Stable Nitrogen Isotopes in parts per million") + geom_point()
print(plot1)

```

The trend is Figure 1 suggests that the stable nitrogen isotopes across polar
bears have remained more or less consistent over the ~15 year period.


```{r,  message = FALSE, echo=FALSE, warning=FALSE}

# Plot 2 - Correlating Time and Percent Nitrogen

plot2 <- ggplot(PolarBearHair, aes(x = Year, y = Percent.N, color = Sex)) + geom_smooth(method=lm) + scale_x_continuous(limits = c(2004, 2020), breaks = seq(1985, 2020, by = 5)) + labs(title = "Figure 2 - Percent of Nitrogen in Polar Bear Hair (2004-2018)",
x = "Year", y = "Percent of Nitrogen by mass measured in the hair") + geom_point()
print(plot2)

```

The trend is Figure 2 suggests that the percentage of nitrogen found in polar
bear hair samples has remained more or less consistent over the ~15 year period. 
Since these nitrogen variables represents the trophic level of polar bear's prey, our results indicate that polar bears have not changed the species they feed on.

```{r,  message = FALSE, echo=FALSE, warning=FALSE}

# Plot 3 - Correlating Time and Carbon Isotopes

plot3 <- ggplot(PolarBearHair, aes(x = Year, y = d13C.VPDB, color = Sex)) + geom_smooth(method=lm) + scale_x_continuous(limits = c(2004, 2020), breaks = seq(1985, 2020, by = 5)) + labs(title = "Figure 3 - Stable Carbon Isotopes sampled from Polar Bear Hair (1985-2018)",
x = "Year", y = "Stable Carbon Isotopes in parts per million") + geom_point()
print(plot3)

```

The trend is Figure 3 suggests that the stable carbon isotopes across polar
bears have remained more or less consistent over the ~15 year period.


```{r,  message = FALSE, echo=FALSE, warning=FALSE}

# Plot 4 - Correlating Time and Percentage Carbon

plot4 <- ggplot(PolarBearHair, aes(x = Year, y = Percent.C, color = Sex)) + geom_smooth(method=lm) + scale_x_continuous(limits = c(2004, 2020), breaks = seq(1985, 2020, by = 5)) + labs(title = "Figure 4 - Percent of Carbon in Polar Bear Hair (1985-2018)",
x = "Year", y = "Percent of Carbon by mass measured in the hair") + geom_point()
print(plot4)

```

The trend is Figure 4 suggests that the percentage of carbon found in polar
bear hair samples has remained more or less consistent over the ~15 year period. 
Since these carbon variables represents whether polar bears feed on marine or terrestrial sources, our results indicate that polar bears have not changed their feeding patterns.
